---
title: "Developmental NonCog Figures V4"
author: "Margherita Malanchini"
date: "13/12/2022"
output: html_document
---

#### Figure 1 #### Now merging phenotypic and trivariate cholesky figures on latent phenotypes. 

2 main messages: Panel A.) NC skills are related to achievement beyond general cognitive ability at every developmental stage and 
Panel B.) this link is partly accoounted for by genetic factors that are common to NC skills and achievement beyond the genetics of cognitive abilty.

```{r, include=FALSE}
rm(list = ls())
library(packcircles);library(tidyverse); library(viridis); library(ggsci)

df= read.csv('/Users/margheritamalanchini/Dropbox/2019 Queen Mary/MM papers 2019/MM_AGA_Genomic SEM Achievement/manuscript/Figures manuscript/Figures data/PHENO_Regressions-latent-factors-lavaan.csv', header = T, strip.white = T ,na.strings = NA)


df$l.bound = df$Estimate-(df$SE*1.96)
df$u.bound = df$Estimate+(df$SE*1.96)

df1 <- mutate(df, age = paste("Age", Age, sep = " "))
df1$agef = factor(df1$age, levels = c("Age 16",
                                      "Age 12",
                                      "Age 9",
                                      "Age 7"))

df1$type = factor(df1$type, levels = c("achievement","cognitive ability","achievement-cognitive"))

df1$rater = factor(df1$rater, levels = c("self", "parent", "teacher"))

df1 = df1[df$type %in% c("achievement-cognitive"), ]
df1$measure_f <- recode_factor(df1$measure, Edu = "Education-related NCS", SR
                        = "Self-regulation")


Figure1A = ggplot(df1, aes(x = agef, y = Estimate))+
geom_bar(stat = "identity", width = .5, position = "dodge", fill = "#f3902c", color = "black")+
geom_errorbar(aes(ymin=l.bound, ymax=u.bound), width=0.1, size=0.8, colour = "black", position=position_dodge(.8)) +
theme(legend.position='right',
    legend.text = element_text(size = 12),
    legend.title = element_text(size = 12, face = "bold"),
    panel.border = element_rect(linetype = 'solid', colour = 'gray',fill = NA),
    panel.background = element_rect(fill = "white",colour = "white",size = 0.5, linetype = "solid"),
    plot.title = element_text(size = 12, 
                              #hjust = 0.5, 
                              face="bold"),
    panel.grid.major = element_line(size = 0.25, linetype = 'solid', colour = "gray"), 
    panel.grid.minor = element_line(size = 0.25, linetype = 'solid'),
    axis.line.x = element_line(colour = 'gray', size=0.5, linetype='solid'),
    axis.line.y = element_line(colour = 'gray', size=0.5, linetype='solid'),
    axis.text.y = element_text(size = 10, colour="black", family = "sans", angle = 0, hjust =0), 
    axis.text.x = element_text(size = 10, colour="black", family = "sans", angle = 0, hjust = 0.5),
    panel.spacing=unit(1.5, "lines"))+
  ylim(0,.75)+ 
  ggtitle("Associations bewteen noncognitive skills and academic \n achievement accounting for general cognitive ability")+
  xlab(" ")+ ylab("Standardized beta coefficient")+
   coord_flip()
  #scale_color_nejm()+
  #scale_fill_nejm()

Figure1A = Figure1A+facet_grid(rater~measure_f,scales="free", space="free")+
  #labs(fill= "Rater:")+
     theme(strip.text.x = element_text(size=12, face="bold"),
      strip.text.y = element_text(size=12, face="bold"),
      strip.background = element_rect(colour="gray", fill="white"))
plot(Figure1A)
#ggsave(Figure1, file="/Users/margheritamalanchini/Dropbox/2019 Queen Mary/MM papers 2019/MM_AGA_Genomic SEM Achievement/manuscript/Figures manuscript/Figures_V2/Figure1.tiff", width=7, height=7) 

```


Figure 1B: Latent trivariate twin models results: change visualize only the genetic portion of the models

```{r, include = FALSE}

df<- read.csv('/Users/margheritamalanchini/Dropbox/2019 Queen Mary/MM papers 2019/MM_AGA_Genomic SEM Achievement/manuscript/Figures manuscript/Figures data/TWIN_Trivariate_latent.csv', header = T, strip.white = T, na.strings = "NA")


scale_fill_ACE_triv <- function(...){
      ggplot2::discrete_scale(
        "fill","ACE_triv", 
        scales::manual_pal(
          values = c(
           "#f94144", "#f3902c","#f8c91e",
           "#277da1","#61a5c2","#a9d6e5",
         "#43aa8b", "#95d5b2","#90be6d")),
        ...)}

scale_color_ACE_triv <- function(...){
      ggplot2::discrete_scale(
        "colour","ACE_triv", 
        scales::manual_pal(
          values = c(
         "#f94144", "#f3902c","#f8c91e",
           "#277da1","#61a5c2","#a9d6e5",
         "#43aa8b", "#95d5b2","#90be6d")),
        ...)}

df$chol_path <- factor(df$chol_path, levels=c( "a31", "a32", "a33", "c31", "c32", "c33",
                                                  "e31", "e32", "e33")) 
# 
df$label <- factor(df$label, levels=c("E ach-noncog-cog", "E noncog-cog","E cog",
                                       "C ach-noncog-cog", "C noncog-cog","C cog",
                                       "A ach-noncog-cog", "A noncog-cog","A cog"
                                       )) 

df <- mutate(df, agef = paste("Age",Age., sep = " "))
df$agef<- factor (df$agef, levels = c( "Age 16",  "Age 12","Age 9", "Age 7"))

df$rater<- factor (df$reporter, levels = c("self", "parent", "teacher"))
df1 = df[df$label %in% c("A ach-noncog-cog", "A noncog-cog","A cog"), ]
#df2$chol_path_r <-fct_rev(df2$chol_path)
#df2$proportion.rp<- as.numeric(paste(df1$proportion.rp))

Figure1B = ggplot(df1,aes(x = agef, y= squared_pe, fill  = label)) +
        geom_bar(position = "stack" ,stat = "identity", color = "black", width = .5) +
        #geom_errorbar(aes(ymin  = l.bound, ymax  = u.bound),width = 0.2,size  = 0.7,
                    #position = pd, color = "black") +
        theme(legend.direction = "horizontal",
              legend.position='none',
              legend.justification = "left",
              legend.text = element_text(size = 12),
              legend.title = element_blank(),
              plot.title = element_text(size = 12, 
                                        #hjust = 0.5, 
                                        face="bold"),
              panel.border = element_rect(linetype = 'solid', colour = 'gray',fill = NA),
              panel.background = element_rect(fill = "white",
                                colour = "white",
                                size = 0.5, linetype = "solid"),
              panel.grid.major = element_line(size = 0.25, linetype = 'solid',
                                colour = "gray"), 
              panel.grid.minor = element_line(size = 0.25, linetype = 'solid'),
              axis.line.x = element_line(colour = 'gray', size=0.5, linetype='solid'),
              axis.line.y = element_line(colour = 'gray', size=0.5, linetype='solid'),
              axis.text.y = element_text(size = 10, colour="black", family = "sans", angle = 0, hjust=0), 
              axis.text.x = element_text(size = 10, colour="black", family = "sans", angle = 0, hjust=0.5), 
              axis.ticks.y = element_blank(),
              axis.title= element_text(size = 12))+
  ylab("Genetic variance in academic achievement")+ 
  ggtitle("Genetic overlap bewteen academic achievement and noncognitive \n skills accounting for general cognitive ability")+ 
  xlab(" ") +
    ylim(0,1)+
    coord_flip()+
  scale_fill_ACE_triv()

  #guides(fill = guide_legend(reverse=TRUE))+
Figure1B = Figure1B+facet_grid(rater~nc_measure2, scales="free", space="free")+ 
    theme (strip.text.x = element_text(size=10, face = "bold"),
                                strip.text.y = element_text(size=10, face="bold"),
                                strip.background = element_rect(colour="gray", fill="white"))

plot(Figure1B)

#ggsave(triv, file="/Users/margheritamalanchini/Dropbox/2019 Queen Mary/MM papers 2019/MM_AGA_Genomic SEM Achievement/manuscript/Figures manuscript/Figures_V2/Figure2.tiff", width=9, height=7) 
```
Bring together Panel A and Panel B 
```{r}
library(gridExtra); library(ggpubr); library (cowplot)

f1<- ggarrange(Figure1A, Figure1B, nrow = 1, ncol = 2)
plot(f1)

#OR

Fig1 = ggdraw() +
draw_plot(Figure1A, .01, .07, .44, .90)+
draw_plot(Figure1B, .45, .07, .49, .90)+
draw_plot_label(c("A", "B"), c(0.02, 0.45), c(.99, .99), size = 14)
Fig1

ggsave(Fig1, file="/Users/margheritamalanchini/Dropbox/2019 Queen Mary/MM papers 2019/MM_AGA_Genomic SEM Achievement/manuscript/Figures manuscript/Figures_V2/Figure1_V2.tiff", width=13, height=8) 


```

Figure 2: Aim provide a summary of our core results in terms of:
Panel A: Genomic SEM model extension
Panel B: Cog Noncog PGS prediction of achievement over developemnt
Panel C: Between/Wihtin family analyses 
Panel D: GxSES analyses 


```{r, include=FALSE}

rm(list = ls())

library(tidyverse); library(gridExtra);library(reshape2)
library(ggsci); library(dplyr); library(psych);
library(forcats)

df <- read.csv('/Users/margheritamalanchini/Dropbox/2019 Queen Mary/MM papers 2019/MM_AGA_Genomic SEM Achievement/manuscript/Figures manuscript/Figures data/PGS_original+extension_for plotting.csv', header = T, stringsAsFactors = F, strip.white = T, na.strings = NA)

# df <- read.csv('/Users/margherita/Dropbox/2019 Queen Mary/MM papers 2019/MM_AGA_Genomic SEM Achievement/manuscript/Figures manuscript/Figures data/PGS_original+extension_for plotting.csv', header = T, stringsAsFactors = F, strip.white = T, na.strings = NA)

#---------------------
df$pheno_category
df$dv_type
#--------------------
#select latent variables for main ms figures and then plot observed variables for SOM
df$dv_type_f = as.factor(df$dv_type)

#Panel B achievement 
df1 = df[df$pheno_category=="Achievement", ]
df1 = df1[df1$phenotype=="Achievement tot", ]
df1$pheno_category_f = as.factor(df1$pheno_category)
df1$age_f = as.factor(df1$age)
df1 = df1[df1$PGS !="Cog", ]
df1 = df1[df1$PGS !="NonCog", ]

b.Ach = df1 %>%
ggplot(aes(x = fct_rev(age_f), y = coeff, ymin=coeff-Robust_SE, ymax=coeff+Robust_SE,color=factor(PGS)))+
geom_pointrange(size = 1.2, position = position_dodge(width = .5))+
theme(legend.direction = "horizontal",
  legend.position="bottom",
        legend.text = element_text(size = 12),
        legend.title = element_text(size = 12),
        panel.border = element_rect(linetype = 'solid', colour = 'gray',fill = NA),
        panel.background = element_rect(fill = "white",
                                        colour = "white",
                                        size = 0.5, linetype = "solid"),
        panel.grid.major = element_line(size = 0.25, linetype = 'solid',
                                        colour = "gray"), 
        panel.grid.minor = element_line(size = 0.25, linetype = 'solid'),
        axis.line.x = element_line(colour = 'gray', size=0.5, linetype='solid'),
        axis.line.y = element_line(colour = 'gray', size=0.5, linetype='solid'),
        axis.text.y = element_text(size = 12, colour="black", family = "sans", angle = 0), 
        axis.text.x = element_text(size = 12, colour="black", family = "sans", angle = 0), 
        axis.ticks.y = element_blank(),
        axis.title= element_text(size = 12),
       plot.title =element_text(size=12, hjust = 0.5))+
  xlab(" ")+
  ylab("Beta and standard error")+
  ggtitle("PGS prediction of acdemic achievement over development")+
  ylim(0, 0.35)+
  geom_hline(yintercept=0, linetype = "dashed", color = "black", size = 0.4)+
  coord_flip()

 b.Ach =  b.Ach +scale_color_manual(values =c( "#000080","#DC143C"),
                               name = "PGS", labels = c( 'CogExt',"NonCogExt"))
  
 b.Ach
   
#ggsave(b.Ach, file="../Figures/tiff/Figure_PGS_Ach.tiff", width=11, height=5)
#ggsave(b.Ach, file="../Figures/png/Figure_PGS_Ach.png", width=11, height=5)

#####################################################################################
#Panel C ---Within-between family effects and GXSES plots


load("/Users/margheritamalanchini/Dropbox/2019 Queen Mary/MM papers 2019/MM_AGA_Genomic SEM Achievement/GSEM/EFA/data_x_DZ/table_results_DZext_prelim_30621.RData")
#load("/Users/margherita/Dropbox/2019 Queen Mary/MM papers 2019/MM_AGA_Genomic SEM Achievement/GSEM/EFA/data_x_DZ/table_results_DZext_prelim_30621.RData")

df = as.data.frame(reslts)
df$depvar_f <- recode_factor(df$depvar, gta1 = "Age 7 ",it2a1 = "Age 9 ", lt2a1 = "Age 12", pcexgcsecorem1 =  "Age 16")

p <- df %>%
  #filter(type == "within") %>%
ggplot(aes(y=fct_rev(depvar_f),x=Estimate, color = PRS, group = PRS)) +
  geom_point(size = 2.5, alpha = 1) +
    geom_path(size = 1, alpha = 1) +
  xlim(0,.4) +
  theme(legend.direction = "vertical",
        legend.position="none",
        legend.text = element_text(size = 12),
        legend.title = element_text(size = 12),
        panel.border = element_rect(linetype = 'solid', colour = 'gray',fill = NA),
        panel.background = element_rect(fill = "white",
                                        colour = "white",
                                        size = 0.5, linetype = "solid"),
        panel.grid.major = element_line(size = 0.25, linetype = 'solid',
                                        colour = "gray"), 
        panel.grid.minor = element_line(size = 0.25, linetype = 'solid'),
        axis.line.x = element_line(colour = 'gray', size=0.5, linetype='solid'),
        axis.line.y = element_line(colour = 'gray', size=0.5, linetype='solid'),
        axis.text.y = element_text(size = 12, colour="black", family = "sans", angle = 0), 
        axis.text.x = element_text(size = 12, colour="black", family = "sans", angle = 0), 
        axis.ticks.y = element_blank(),
        axis.title= element_text(size = 12))+
  xlab("Beta and standard error")+
  ylab(" ")+
  facet_grid(~type)+ theme (strip.text.x = element_text(size=10, face = "bold"),
                                strip.text.y = element_text(size=10, face="bold"),
                                strip.background = element_rect(colour="gray", fill="white"))
p= p +scale_color_manual(values =c( "#000080","#DC143C"),
                               name = "PGS",
                          labels = c( 'Cog ext',"NonCog ext"))
p

#ggsave(p, file="../Figures/tiff/withinbewteenGSEM.tiff", width=7, height=6)
#ggsave(p, file="../Figures/png/withinbewteenGSEM.png", width=7, height=6)

#############################################################################
#Panel D PGSxSES interaction

vars <- c("gta1","it2a1","lt2a1","pcexgcsecorem1") 

setwd('/Users/margherita/Dropbox/2019 Queen Mary/MM papers 2019/MM_AGA_Genomic SEM Achievement/GSEM/EFA')

df <- read.csv('/Users/margherita/Dropbox/2019 Queen Mary/MM papers 2019/MM_AGA_Genomic SEM Achievement/GSEM/EFA/2021-04-26_DevNoncog_prepared-v4-latentfactors.csv')

df$NonCog <- df$NonCogFRC_1
df$Cog <- df$CogFRC_1
df$CogExt <- df$c_chol_cog_noncog_FRC_1
df$NonCogExt <- df$nc_chol_cog_noncog_FRC_1


resultsCog <- list()
resultsNonCog <- list()

for (i in vars) { 
  
KeepVars = c(i,'NonCog','Cog','PC1','PC2','PC3','PC4','PC5','PC6','PC7','PC8','PC9','PC10','Chip','Batch','sex1')

#dat = for conditional plot 
dat <- df[complete.cases(df[,KeepVars]), ]

qSES <- quantile(dat$ases, na.rm = T) 

dat$sesCoars<- ifelse(dat$ases< qSES[2], "LOW SES", #lowest quantile
       ifelse(dat$ases> qSES[4],"HIGH SES", #highest quantile
              "MEDIUM SES")) #middle 50%

dat <- dat %>%
  filter(!is.na(sesCoars))

dat$sesCoars = factor(dat$sesCoars, levels=c("LOW SES","MEDIUM SES","HIGH SES"), labels=c("LOW SES","MEDIUM SES","HIGH SES")) 

#jointModelling
covars = c('PC1','PC2','PC3','PC4','PC5','PC6','PC7','PC8','PC9','PC10','Chip','Batch','sex1')

indep <- c("ases","NonCog","Cog") #test interactions separately for cog non cog 
int <- apply(expand.grid(indep, covars), 1, function(x) paste(x[2], x[1], sep=":")) 

fCog <- as.formula(paste(i,"~", paste(indep, collapse = '+'),'+', paste(int, collapse = '+'),'+',paste(covars, collapse = '+'),'+  ases:Cog'))
resultsCog[[paste0("joint",i)]] <- summary(do.call("lm", list(fCog, data = quote(dat)))) #list of summary lm()

fNCog <- as.formula(paste(i,"~", paste(indep, collapse = '+'),'+', paste(int, collapse = '+'),'+',paste(covars, collapse = '+'),'+  ases:NonCog'))
resultsNonCog[[paste0("joint",i)]] <- summary(do.call("lm", list(fNCog, data = quote(dat)))) #list of summary lm()

### std residual loop
for (ind in c("Cog", "NonCog")){
f <- as.formula(paste0(i, "~", ind,"+",paste(covars, collapse = '+')))
dat[,paste0("rs",i,ind)] <- rstandard(do.call("lm", list(f, quote(dat))))
}

yCog <- paste0("rs",i,"Cog")
yNCog <- paste0("rs",i,"NonCog")


x <- ggplot(dat) + 
  geom_jitter(aes_string("NonCog",noquote(yCog)), colour="#DC143C",alpha=0.5) + 
  geom_smooth(aes_string("NonCog",noquote(yCog)), method=lm, colour = "#DC143C", se=FALSE) +
  geom_jitter(aes_string("Cog",noquote(yNCog)), colour="#000080",alpha=0.3) + 
  geom_smooth(aes_string("Cog",noquote(yNCog)), method=lm, colour = '#000080',  se=FALSE) +
  geom_boxplot(aes_string("Cog", y = -4.5), size = .2
               , width = .2
               ,colour="#000080"
               ,alpha=0.4
               ) +
  geom_boxplot(aes_string(x = "NonCog", y = -5), size = .2
                , width = .2
               ,colour="#DC143C"
               ,alpha=0.5
               ) +
  theme(panel.border = element_rect(linetype = 'solid', colour = 'gray',fill = NA),
        panel.background = element_rect(fill = "white",
                                        colour = "white",
                                        size = 0.5, linetype = "solid"),
    panel.grid.major = element_line(size = 0.25, linetype = 'solid',
                                        colour = "gray"), 
        panel.grid.minor = element_line(size = 0.25, linetype = 'solid'),
        axis.line.x = element_line(colour = 'gray', size=0.5, linetype='solid'),
        axis.line.y = element_line(colour = 'gray', size=0.5, linetype='solid'),
        axis.text.y = element_text(size = 12, colour="black", family = "sans", angle = 0), 
        axis.text.x = element_text(size = 12, colour="black", family = "sans", angle = 0), 
        axis.ticks.y = element_blank(),
        axis.title= element_text(size = 12))+
  geom_hline(yintercept = 0, linetype = "dashed", colour = "black", alpha = .3) +
   geom_vline(xintercept = 0, linetype = "dashed", colour = "black", alpha = .3) + 
  facet_wrap(~sesCoars, scales="free_x") +
  labs(x = "PRS", y = i)  

png(paste0(i,"_adj.png"), res=300, height = 1200, width = 2000)
  plot(x)
dev.off()

}

Achgxses = x
Achgxses = Achgxses+labs(x = "PGS", y = "Academic achievement age 16")+ theme(strip.text.x = element_text(size=10, face = "bold"),
                                strip.text.y = element_text(size=10, face="bold"),
                                strip.background = element_rect(colour="gray", fill="white"))

Achgxses

library(cowplot)
require(magick)

leg <- get_legend(b.Ach)
b.Ach = b.Ach + theme(legend.position = "none")

Fig2 = ggdraw() +
draw_image("/Users/margherita/Dropbox/2019 Queen Mary/MM papers 2019/MM_AGA_Genomic SEM Achievement/manuscript/Figures manuscript/Figures_V2/NonCogExt Model.tiff", 0, .5, .45, .42)+
draw_plot(b.Ach, .5, .5, .43, .42)+
draw_plot(p, .01, .05, .35, .42)+
draw_plot(Achgxses,.40, .05, .57,.42)+ 
draw_plot(leg, .1, .00, .90, .05)+
draw_plot_label(c("A", "B", "C", "D"), c(0.02, 0.5, 0.02, 0.40), c(.95, .95, .50, .50), size = 14)
Fig2

ggsave(Fig2, file="/Users/margherita/Dropbox/2019 Queen Mary/MM papers 2019/MM_AGA_Genomic SEM Achievement/manuscript/Figures manuscript/Figures_V2/Figure2new.tiff", width=11, height=8) 

ggsave(Fig2, file="/Users/margherita/Dropbox/2019 Queen Mary/MM papers 2019/MM_AGA_Genomic SEM Achievement/manuscript/Figures manuscript/Figures_V2/Figure2new.png", width=11, height=8) 

```


